---
title: "Postprocessing PopRep Results"
author: "Peter von Rohr"
date: "10/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Disclaimer
The processing steps in GenMon with the PopRep results are described in this document.


## Basic Concept
In the top section of the script 'PopRep.php', the PopRep program is called using the 'exec()' interface of php to the bash shell. PopRep is started with the bash-script 'process_uploads.sh'. This runs the whole PopRep pipeline on the uploaded extended pedigree. The results of PopRep are stored in the PopRep database. The name of the database is the same as the name of the PopRep project that was created for the analysis of the uploaded pedigree. 

The results of PopRep that are used by GenMon are extracted from the PopRep-Project database into the database for GenMon called 'GenMon_CH'. After the transfer of the PopRep-results into the 'GenMon_CH' database, the results are post-processed. In what follows, we try to understand these postprocessing steps.


### Running php-script
All computations including the PopReport analysis and all post-processing steps are done from inside of 'PopRep.php'. This script is either executed when pushing the 'Add'-button on the breed-detail page. 

According to 'https://alvinalexander.com/php/how-to-run-php-script-from-linux-unix-command-line/' and 'https://stackoverflow.com/questions/6965541/run-php-code-in-command-line', the php-script can also be run via the commandline tool called 'php'. 

```
Singularity> (echo '<?php';echo 'echo "Hello "; echo "Jay\n";';echo '?>') > hello.php
Singularity> php -f hello.php 
```


## Project
The first step of the post-processing consists of retrieving the project name from the postgresql system table 'pg_database'. The project name is stored in the php-variable '$project_name'. Based on the 'sql'-statement, it is assumed that only one project with a name starting with 'PPP-' is in the table 'pg_database'. What happens when several users try to run an analysis is not clear at this point.


## Table Rename
Certain PopRep tables are renamed to names that contain the breed ID or name. The following relation between old and new tables is taken from 'PopRep.php'.

```{r table-rename, echo=FALSE, results='asis'}
tbl_rnm <- tibble::tibble(`New Table` = c('"breed".$breed_id."_inbryear"',
                                          '"breed".$breed_id."_inb_year_sex"',
                                          '"breed".$breed_id."_ne"',
                                          '"breed".$breed_id."_pedcompl"',
                                          '"breed".$breed_id."_ne_deltaF"',
                                          '"breed".$breed_id."_transfer"',
                                          '"breed".$breed_id."_data"',
                                          'gene_stuff',
                                          'tmp1_gen'),
                          `Old Table` = c('tmp2_table3',
                                          'tmp2_table2',
                                          'tmp2_ne',
                                          'tmp2_pedcompl',
                                          'tmp2_table5',
                                          'transfer',
                                          'animal',
                                          'gene_stuff',
                                          'tmp1_gen'))
knitr::kable(tbl_rnm)
```

The renaming process is done in the `while`-loop between lines 43-71. The first steps in the loop drop any existing tables. After the dropping of the tables, the old tables from the PopRep-database are dumped using the command-line-tool `pg_dump`. The output of `pg_dump` consist of sql-statements which are piped into the `psql` client which imports the dumped data into the GenMon database 'GenMon_CH'. The 'SCHEMA' of the newly imported data into the database 'GenMon_CH' is changed from 'apiis' to 'public'. Whenever the rename is not successful, the PopRep results are removed.

The following bash command can be used to check whether the re-naming was successful.

```
psql -U geome_admin -d GenMon_CH -c "select * from pg_catalog.pg_tables where tableowner = 'apiis_admin';"
```

The check shows that we cannot find the table 'breed6_data' which should be the result of dumping and importing the table 'animal'. The problem is debugged by running 'PopRep' alone and by checking the availablility of the different tables. The following command runs PopRep

```
./simg/genmon-sidef/bash/run_poprep.sh -b ctb20201104 -p /var/www/html/genmon-ch/pdf/data_sample.csv -u www-data -g www-data -i /var/lib/postgresql/incoming -l /home/quagadmin/gnm/gnmlog/popreport.log
```

The script 'PopRep.php' selects the database name with all the PopRep results with the following statement.

```
psql -U postgres -d GenMon_CH -c "SELECT datname FROM pg_database where datname like 'PPP%';"
```


After running poprep the following command shows the tables

```
PRPPROJ=PPP_hWfogGKhaTwlU
psql -U postgres -d $PRPPROJ -c "select * from pg_catalog.pg_tables where tableowner = 'apiis_admin';"
```

The following command shows that table 'animal' does exist.

```
PRPPROJ=PPP_hWfogGKhaTwlU
psql -U postgres -d $PRPPROJ -c "select * from pg_catalog.pg_tables where tableowner = 'apiis_admin' and tablename = 'animal';"
```



Between lines 98 to 111 indices of certain tables are renamed. 


## Re-coding Sexes
The sex codes are changed in the table 'breed6_data'. 

## Check Content of Table `breed` 

```
psql -U postgres -d GenMon_CH -c "select * from breed6_data;"
```

Any table with a name starting with 'breed'

```
psql -U postgres -d GenMon_CH -c "select * from pg_catalog.pg_tables where tablename like 'breed_%';"
```

Tables without a breed-ID

```
psql -U postgres -d GenMon_CH -c "select * from pg_catalog.pg_tables where substring(tablename from 1 for 6) = 'breed_';"
```

Drop the tables without a breed_ID

```
psql -U postgres -d GenMon_CH -c "drop table breed_pedcompl, breed_ne_deltaf, breed_inb_year_sex, breed_transfer, breed_ne, breed_data;"
```

Drop tables for breed-ID 6

```
psql -U postgres -d GenMon_CH -c "drop table breed6_inbryear, breed6_pedcompl, breed6_inb_year_sex, breed6_inb_plz, breed6_transfer, breed6_ne, breed6_ne_deltaf;"
```

```
psql -U geome_admin -d GenMon_CH -c "select * from pg_catalog.pg_tables where tablename like 'breed_%';"
```


# Stepwise Post-Processing 
## Step 1: Renaming Tables and Indices

What are the tables after the renaming process

```
psql -U geome_admin -d GenMon_CH -c "select * from pg_catalog.pg_tables;"
```

All tables related to 'breed6'

```
psql -U geome_admin -d GenMon_CH -c "select * from pg_catalog.pg_tables where tablename like 'breed6%';"
```

The table 'breed6_data'

```
psql -U geome_admin -d GenMon_CH -c "select * from breed6_data;"
```

Describe the table

```
psql -U geome_admin -d GenMon_CH -c "select column_name, data_type, character_maximum_length, column_default, is_nullable
from INFORMATION_SCHEMA.COLUMNS where table_name = 'breed6_data';"
```

This shows that columns like `plz` are text and hence cannot be used with an integer in a where condition.


## Step 2: Updating sex

```
psql -U geome_admin -d GenMon_CH -c "select * from breed6_data where db_sex = 117;"
```

Distinct values of `db_sex`

```
psql -U geome_admin -d GenMon_CH -c "select distinct db_sex from breed6_data;"
```

Run the update-statement

```
psql -U geome_admin -d GenMon_CH -c "UPDATE breed6_data SET db_sex=2 where db_sex=117;"
psql -U geome_admin -d GenMon_CH -c "UPDATE breed6_data SET db_sex=3 where db_sex=118;"
psql -U geome_admin -d GenMon_CH -c "select distinct db_sex from breed6_data;"
```

Frequencies

```
psql -U geome_admin -d GenMon_CH -c "select db_sex, count(db_sex) as sex_counts from breed6_data group by db_sex order by db_sex;"
```

SELECT 
  traveler_answers.choice, 
  COUNT(traveler_answers.choice) as total_answers
FROM traveler_answers
GROUP BY traveler_answers.choice
ORDER BY traveler_answers.choice 

## Step 3: Update MVC

```
# columns: 'plz', 'introgression', 'inb_gen', 'cryo'
# plz
psql -U geome_admin -d GenMon_CH -c "update breed6_data set plz=NULL where plz='-9999';"
psql -U geome_admin -d GenMon_CH -c "select plz, count(plz) as plz_counts from breed6_data group by plz order by plz;"
# introgression
psql -U geome_admin -d GenMon_CH -c "update breed6_data set introgression=NULL where introgression='-9999';"
psql -U geome_admin -d GenMon_CH -c "select introgression, count(introgression) as introgression_counts from breed6_data group by introgression order by introgression;"
# inb_gen
psql -U geome_admin -d GenMon_CH -c "update breed6_data set inb_gen=NULL where inb_gen='-9999';"
psql -U geome_admin -d GenMon_CH -c "select inb_gen, count(inb_gen) as inb_gen_counts from breed6_data group by inb_gen order by inb_gen;"
# cryo
psql -U geome_admin -d GenMon_CH -c "update breed6_data set cryo_cons=NULL where cryo_cons='-9999';"
psql -U geome_admin -d GenMon_CH -c "select cryo_cons, count(cryo_cons) as cryo_cons_counts from breed6_data group by cryo_cons order by cryo_cons;"

```


## Step 4: Inbreeding

```
psql -U geome_admin -d GenMon_CH -c "select * from breed6_ne;"
```

```
psql -U geome_admin -d GenMon_CH -c "select * from breed6_ne_deltaF;"
```

```
psql -U geome_admin -d GenMon_CH -c "select * from tmp1_gen;"
```

Check whether inbreeding column is added

```
psql -U geome_admin -d GenMon_CH -c "select * from breed6_data;"
# table description
psql -U geome_admin -d GenMon_CH -c "select column_name, data_type, character_maximum_length, column_default, is_nullable
from INFORMATION_SCHEMA.COLUMNS where table_name = 'breed6_data';"
```


## Step 5: Further Steps - Type Problems

```
psql -U geome_admin -d GenMon_CH -c "select column_name, data_type, character_maximum_length, column_default, is_nullable
from INFORMATION_SCHEMA.COLUMNS where table_name = 'breed6_inb_plz';"
```

In all tables except 'breed6_data' the plz column is of type integer, hence try a cast

```
psql -U geome_admin -d GenMon_CH -c "alter table breed6_data alter column plz TYPE INTEGER  USING (plz::integer);"
psql -U geome_admin -d GenMon_CH -c "select column_name, data_type, character_maximum_length, column_default, is_nullable
from INFORMATION_SCHEMA.COLUMNS where table_name = 'breed6_data';"
```

More type missmatch

```
PHP Warning:  pg_query(): Query failed: ERROR:  function avg(text) does not exist
LINE 4: (select avg(bd.inb_gen) as in, bd.plz as p
```

caused by the following statement

```
UPDATE breed".$breed_id."_inb_plz
SET mean_inb_gen_lastgi =
(select q.in from
(select avg(bd.inb_gen) as in, bd.plz as p
from breed".$breed_id."_data bd
where extract(year from bd.birth_dt)>=(".$max_year."-".$GI.")
group by bd.plz) q
where q.p=breed".$breed_id."_inb_plz.plz)
```

Unroll





## Problems, TODOS and FAQ

* `[FAQ]`: What happens with project selection when more than one user is using the system?


* `[Problem 1]`: In `FunctionsCalcIndex.php` the computations in function `Min_radius()` can run in an infinite loop. The problem is that the possibility of the `while`-condition not being met is not catched. Furthermore the division in the `while`-loop to compute the percentage can produce a division-by-zero exception. 
    + `[Solution 1]`: Two `if`-statements are introduced to avoid the division-by-zero exception and the infinite loop.
    
    
* `[Problem 2]`: In `PopRep.php` the database of the PopRep-project is selected form the system table 'pg_database' using the following sql-statement: `SELECT datname FROM pg_database where datname like 'PPP%';` This means that if more than one analysis is running concurrently, it is possible that the wrong PopRep-project database is chosen. 

