---
title: "Installation of Genmon"
author: "Peter von Rohr"
date: "8/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Disclaimer
Installation of Genmon in Singularity Container. The installation includes PopRep for the analysis of pedigrees and all programs for running the GIS and the socio-eco analyses. 


## Goal
The complete set of software programs are to be installed inside of a container. 


## Strategy
Following the approach taken with the installation of PopRep in a container, we want to extend the existing definition file for the poprep container with the missing parts to run the GIS and the Socio-Eco analyses. 

The installation steps follow the instructions given by https://github.com/SolangeD/GENMON/blob/master/install.html.


## Step-by-Step Protocol
This section contains a protocol showing all steps to build a container running all components of GenMon

1. The definition file 'Singularity_poprep-recipe' is taken as a base for the new container image
2. The parts not included by the definition file of the poprep container are added to the definition file 'Singularity_genmon.recipe'. 


## Check List based on `install.html`
The following table gives an overview about the installation status of the different points listed on `install.html`.

```{r, echo=FALSE}
tbl_install_status <- tibble::tibble(Section = c("OS Install",
                                                 "Server Basics",
                                                 "Firewall",
                                                 "Postgres and Postgis",
                                                 "Postgres and Postgis",
                                                 "Install PHP",
                                                 "Install PHP",
                                                 "Install GeoServer",
                                                 "Install GeoServer",
                                                 "Install GENMON",
                                                 "Install PopRep"),
                                     Command = c("",
                                                 "apt-get install vim build-essential",
                                                 "ufw ...",
                                                 "apt-get install postgresql-10-postgis-2.4",
                                                 "import DB ...",
                                                 "apt-get install php5-pgsql",
                                                 "Change config",
                                                 "curl <geoserver-uri> ;unzip ...;mv",
                                                 "jdk;tomcat",
                                                 "...",
                                                 "..."),
                                     Status  = c("simg def",
                                                 "simg def",
                                                 "host",
                                                 "simg def",
                                                 NA,
                                                 "simg def",
                                                 NA,
                                                 "simg def",
                                                 "simg def",
                                                 NA,
                                                 "simg def"))
knitr::kable(tbl_install_status)
```


