---
title: "Installation of Genmon"
author: "Peter von Rohr"
date: "8/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Disclaimer
Installation of Genmon in Singularity Container. The installation includes PopRep for the analysis of pedigrees and all programs for running the GIS and the socio-eco analyses. 


## Goal
The complete set of software programs are to be installed inside of a container. 


## Strategy
Following the approach taken with the installation of PopRep in a container, we want to extend the existing definition file for the poprep container with the missing parts to run the GIS and the Socio-Eco analyses. 

The installation steps follow the instructions given by https://github.com/SolangeD/GENMON/blob/master/install.html.


## Step-by-Step Protocol
This section contains a protocol showing all steps to build a container running all components of GenMon

1. The definition file 'Singularity_poprep-recipe' is taken as a base for the new container image
2. The parts not included by the definition file of the poprep container are added to the definition file 'Singularity_genmon.recipe'. 


## Check List based on `install.html`
The following table gives an overview about the installation status of the different points listed on `install.html`.

```{r, echo=FALSE}
tbl_install_status <- tibble::tibble(Section = c("OS Install",
                                                 "Server Basics",
                                                 "Firewall",
                                                 "Postgres and Postgis",
                                                 "Postgres and Postgis",
                                                 "Install PHP",
                                                 "Install PHP",
                                                 "Install GeoServer",
                                                 "Install GeoServer",
                                                 "Install GENMON",
                                                 "Install PopRep"),
                                     Command = c("",
                                                 "apt-get install vim build-essential",
                                                 "ufw ...",
                                                 "apt-get install postgresql-10-postgis-2.4",
                                                 "import DB ...",
                                                 "apt-get install php5-pgsql",
                                                 "Change config",
                                                 "curl <geoserver-uri> ;unzip ...;mv",
                                                 "jdk;tomcat",
                                                 "...",
                                                 "..."),
                                     Status  = c("simg def",
                                                 "simg def",
                                                 "host",
                                                 "simg def",
                                                 NA,
                                                 "simg def",
                                                 NA,
                                                 "simg def",
                                                 "simg def",
                                                 "simg def",
                                                 "simg def"))
knitr::kable(tbl_install_status)
```


## Open Issues
This section lists all problems and open issues after the first installation step which is implemented in the singularity definition file. 

### Database
The database must be configured and the database server must be started. Once the database is running, the given file 'empty_GenMon_DB.sql' must be read into the database. The required directories for the data and the logfiles are created by the script 'init_gnm_workdir.sh'. 

The import of the database dump is included in 'post_install_gnm.sh'. When reading 'empty_GenMon_DB.sql' the following error comes up

```
ERROR:  relation "public.ofs_" does not exist
```

The reason for this seams to be that the owner of the table 'public.ofs_' is tried to be changed. But this table is not created before. Hence this can be ignored. In the post-installation script 'post_install_gnm.sh' the statement is commented out before reading the database dump.

### Webserver
We have to find out how to run the apache2 webserver in a container. Based on 'https://stackoverflow.com/questions/47297645/binding-ports-when-running-docker-images-in-singularity' there is not network isolation in singularity, at least in older versions. According to 'https://sylabs.io/guides/3.0/user-guide/networking.html' a feature called 'Network virtualisation' can be used in versions 3.0 and later of singularity.

An alternative way is to install nginx and apache2 side-by-side. The setup is such that nginx is the primary server and the requests for a specific port is answered by apache2. This setup is done according to the description under 'https://www.digitalocean.com/community/tutorials/how-to-configure-nginx-as-a-web-server-and-reverse-proxy-for-apache-on-one-ubuntu-18-04-server'. The mentioned description is modified in a way such that the apache2 webserver is running in a singularity container. From what is shown from the logfile nginx does the proxy-redirection, but apache is not responding. We try to solve this by looking into singularity services as described in the next paragraph.

A search for 'singularity hub apache webserver' returned a description of how to run an apache webserver in a container which is available under 'https://github.com/singularityhub/singularity-compose-examples/tree/master/apache-simple#singularity-only'. In that description the arguments '--net' and  '--network-args' specifying a portmap were included in the singularity instance start command. According to the singularity documentation under 'https://sylabs.io/guides/3.0/user-guide/networking.html' this requires v 3.0 and later of the singularity host. Hence we have to upgrade the singularity host on fagr.

### PopRep
The description given in 'install.html' mentions a number of changes to the code of PopRep.  


## Upgrade Singularity
The current version of the singularity container host which comes from the bionic repository is 

```
$ singularity --version
2.6.1-dist
```

We want to upgrade singularity container host to a version past 3.0. The first step to do this is to uninstall the current version. Then we try to install a more current version.

### Step 1: Uninstall
The following command uninstalls the singularity container host

```
sudo apt purge singularity-container
```

We can check with

```
sudo apt list --installed | grep singularity
```

or

```
dpkg --list singularity-container
```

### Step 2: Requirements
In a first step, the requirements listed on '' are checked

```
for p in ca-certificates containernetworking-plugins libc6 libseccomp2 squashfs-tools
do
  echo " * Checking $p ..."
  dpkg --list $p
  sleep 2
done
```

From the output of the above, we have to install 'containernetworking-plugins'. 

```
DWLDIR=/home/quagadmin/source/singularity-container
if [ ! -d "$DWLDIR" ];then mkdir -p $DWLDIR;fi
cd $DWLDIR
wget http://ftp.br.debian.org/debian/pool/main/g/golang-github-containernetworking-plugins/containernetworking-plugins_0.8.6-2_amd64.deb
sudo dpkg --install containernetworking-plugins_0.8.6-2_amd64.deb
#rm -rf containernetworking-plugins_0.8.6-2_amd64.deb
```

### Step 3: Install

The new version is downloaded from 

```
wget http://ftp.br.debian.org/debian/pool/main/s/singularity-container/singularity-container_3.5.2+ds1-1_amd64.deb
sudo dpkg --install singularity-container_3.5.2+ds1-1_amd64.deb
```

After the upgrade the previously created container images could still be run. The only change that had to be done was to add the container directory /var/run/postgresql to the directory binding list.


## Webserver as Service
We use the description under 'https://sylabs.io/guides/3.0/user-guide/running_services.html' for the services and 'https://sylabs.io/guides/3.0/user-guide/networking.html' for the network-arguments as a template. 

The following command pulls and builds the container with an nginx webserver inside.

```
sudo singularity instance start --writable-tmpfs --net --network-args "portmap=8080:80/tcp" docker://nginx web2
```

The container must be started and can then be tested using 'curl'. 

```
sudo singularity exec instance://web2 nginx
$ curl localhost:8080
10.22.0.1 - - [16/Oct/2018:09:34:25 -0400] "GET / HTTP/1.1" 200 612 "-" "curl/7.58.0" "-"
<!DOCTYPE html>
<html>
<head>
<title>Welcome to nginx!</title>
<style>
    body {
        width: 35em;
        margin: 0 auto;
        font-family: Tahoma, Verdana, Arial, sans-serif;
    }
</style>
</head>
<body>
<h1>Welcome to nginx!</h1>
<p>If you see this page, the nginx web server is successfully installed and
working. Further configuration is required.</p>

<p>For online documentation and support please refer to
<a href="http://nginx.org/">nginx.org</a>.<br/>
Commercial support is available at
<a href="http://nginx.com/">nginx.com</a>.</p>

<p><em>Thank you for using nginx.</em></p>
</body>
</html>
```

### Apache 
We use apache2 instead of nginx

```
sudo singularity instance start --writable-tmpfs --net --network-args "portmap=8080:80/tcp" docker://httpd apache2
```

Starting the server

```
sudo singularity exec instance://apache2 httpd
$ curl localhost:8080
<html><body><h1>It works!</h1></body></html>
```


## Integration
The findings from above on how to run an apache2 webserver in a singularity instance are now applied to the singularity image that was created before. 

In a first step, the start of a genmon instance is prepared by running 

```
./simg/genmon-sidef/bash/init_gnm_workdir.sh
```

Then the instance with GenMon can be started as

```
sudo singularity instance start  --writable-tmpfs --net --network-args "portmap=8080:8080/tcp" /home/quagadmin/simg/img/genmon/gnm.simg signm
```

The successful start of the instance can be verified using

```
$ sudo singularity instance list
INSTANCE NAME    PID      IP              IMAGE
signm            22091    10.22.0.5       /home/quagadmin/simg/img/genmon/gnm.simg
```

The apache webserver then has to be started with 

```
sudo singularity exec instance://signm httpd
/.singularity.d/actions/exec: 21: exec: httpd: not found
```

In '/.singularity.d/actions/exec' all bash scripts contained in '/.singularity.d/env' are executed. In the 'apache2' instance there is the script '10-docker2singularity.sh' contains the extension of the PATH variable with the path to httpd.

Found the apache2 executable and trying to call it directly

```
sudo singularity exec instance://signm /usr/sbin/apache2
[Thu Aug 20 15:01:13.761430 2020] [core:warn] [pid 16] AH00111: Config variable ${APACHE_RUN_DIR} is not defined
apache2: Syntax error on line 80 of /etc/apache2/apache2.conf: DefaultRuntimeDir must be a valid directory, absolute or relative to ServerRoot
```

Try with based on information in /etc/apache2/apache2.conf

```
sudo singularity exec instance://signm /usr/sbin/apache2ctl start
```

Checking leads to 

```
$ curl localhost:8080
curl: (7) Failed to connect to localhost port 8080: Connection refused
```

For some reasons the wrong port was specified in ports.conf. With 

```
 1838  sudo singularity instance start --writable-tmpfs --net --network-args "portmap=8080:80/tcp" /home/quagadmin/simg/img/genmon/gnm.simg signm
 1839  sudo singularity exec instance://signm /usr/sbin/apache2ctl start
 1840  curl localhost:8080
```

It works.





## TODOs
This section contains problems and issues which have not been resolved yet.


### Postgresql Port
Check whether the environment variable 'PGPORT' which defines on which port postgresql is listening can be defined in the singularity definition file. According to 'https://sylabs.io/guides/3.0/user-guide/environment_and_metadata.html' this can be done with a special label '%environment' in the singularity definition file. The current running version of 'ssgnm' does not seam to need this. 

