---
title: "Installation of Genmon"
author: "Peter von Rohr"
date: "8/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Disclaimer
Installation of Genmon in Singularity Container. The installation includes PopRep for the analysis of pedigrees and all programs for running the GIS and the socio-eco analyses. 


## Goal
The complete set of software programs are to be installed inside of a container. 


## Strategy
Following the approach taken with the installation of PopRep in a container, we want to extend the existing definition file for the poprep container with the missing parts to run the GIS and the Socio-Eco analyses. 

The installation steps follow the instructions given by https://github.com/SolangeD/GENMON/blob/master/install.html.


## Step-by-Step Protocol
This section contains a protocol showing all steps to build a container running all components of GenMon

1. The definition file 'Singularity_poprep-recipe' is taken as a base for the new container image
2. The parts not included by the definition file of the poprep container are added to the definition file 'Singularity_genmon.recipe'. 


## Check List based on `install.html`
The following table gives an overview about the installation status of the different points listed on `install.html`.

```{r, echo=FALSE}
tbl_install_status <- tibble::tibble(Section = c("OS Install",
                                                 "Server Basics",
                                                 "Firewall",
                                                 "Postgres and Postgis",
                                                 "Postgres and Postgis",
                                                 "Install PHP",
                                                 "Install PHP",
                                                 "Install GeoServer",
                                                 "Install GeoServer",
                                                 "Install GENMON",
                                                 "Install PopRep"),
                                     Command = c("",
                                                 "apt-get install vim build-essential",
                                                 "ufw ...",
                                                 "apt-get install postgresql-10-postgis-2.4",
                                                 "import DB ...",
                                                 "apt-get install php5-pgsql",
                                                 "Change config",
                                                 "curl <geoserver-uri> ;unzip ...;mv",
                                                 "jdk;tomcat",
                                                 "...",
                                                 "..."),
                                     Status  = c("simg def",
                                                 "simg def",
                                                 "host",
                                                 "simg def",
                                                 NA,
                                                 "simg def",
                                                 NA,
                                                 "simg def",
                                                 "simg def",
                                                 "simg def",
                                                 "simg def"))
knitr::kable(tbl_install_status)
```


## Open Issues
This section lists all problems and open issues after the first installation step which is implemented in the singularity definition file. 

### Database
The database must be configured and the database server must be started. Once the database is running, the given file 'empty_GenMon_DB.sql' must be read into the database. The required directories for the data and the logfiles are created by the script 'init_gnm_workdir.sh'. 

The import of the database dump is included in 'post_install_gnm.sh'. When reading 'empty_GenMon_DB.sql' the following error comes up

```
ERROR:  relation "public.ofs_" does not exist
```

The reason for this seams to be that the owner of the table 'public.ofs_' is tried to be changed. But this table is not created before. Hence this can be ignored. In the post-installation script 'post_install_gnm.sh' the statement is commented out before reading the database dump.

### Webserver
We have to find out how to run the apache2 webserver in a container. Based on 'https://stackoverflow.com/questions/47297645/binding-ports-when-running-docker-images-in-singularity' there is not network isolation in singularity, at least in older versions. According to 'https://sylabs.io/guides/3.0/user-guide/networking.html' a feature called 'Network virtualisation' can be used in versions 3.0 and later of singularity.

An alternative way is to install nginx and apache2 side-by-side. The setup is such that nginx is the primary server and the requests for a specific port is answered by apache2. This setup is done according to the description under 'https://www.digitalocean.com/community/tutorials/how-to-configure-nginx-as-a-web-server-and-reverse-proxy-for-apache-on-one-ubuntu-18-04-server'. The mentioned description is modified in a way such that the apache2 webserver is running in a singularity container. From what is shown from the logfile nginx does the proxy-redirection, but apache is not responding. We try to solve this by looking into singularity services as described in the next paragraph.

A search for 'singularity hub apache webserver' returned a description of how to run an apache webserver in a container which is available under 'https://github.com/singularityhub/singularity-compose-examples/tree/master/apache-simple#singularity-only'. In that description the arguments '--net' and  '--network-args' specifying a portmap were included in the singularity instance start command. According to the singularity documentation under 'https://sylabs.io/guides/3.0/user-guide/networking.html' this requires v 3.0 and later of the singularity host. Hence we have to upgrade the singularity host on fagr.

### PopRep
The description given in 'install.html' mentions a number of changes to the code of PopRep.  

### Install GenMon
The section 'Install the GENMON code' in the document 'https://github.com/SolangeD/GENMON/blob/master/install.html' describes the domain-name 'localhost' must be replaced by the actual domain-name of the server on which GenMon is installed. A grep for the term 'localhost' gives the following result

```
grep localhost /var/www/html/genmon-ch/*
/var/www/html/genmon-ch/connectDataBase.php:	$dbh = pg_connect("host=localhost port=5432 dbname=GenMon_CH user=geome_admin password=geome");
grep: /var/www/html/genmon-ch/flot: Is a directory
/var/www/html/genmon-ch/FunctionsGenDiv.php:			OpenLayers.ProxyHost= \"http://localhost/cgi-bin/proxy.cgi?url=\";
/var/www/html/genmon-ch/FunctionsGenDiv.php:               \"http://localhost:8085/geoserver/cite/wms?\",
/var/www/html/genmon-ch/FunctionsGenDiv.php:				sld: 'http://localhost/genomap/".$nameXML.".xml',
/var/www/html/genmon-ch/GenAnimal.php:$_SESSION['hostDirectory'] = "http://localhost/genmon-ch/Data_files/";	// host directory needed by the javaScript script to locate the .json file
/var/www/html/genmon-ch/GenDivPieChartsMapping.php:$_SESSION['hostDirectory'] = "http://localhost/genmon-ch/Data_files/";	// host directory needed by the javaScript script to locate the .json file
/var/www/html/genmon-ch/GenOFS.php:$_SESSION['hostDirectory'] = "http://localhost/genmon-ch/Data_files/";	// host directory needed by the javaScript script to locate the .json file
/var/www/html/genmon-ch/GenStuDb.php:$_SESSION['hostDirectory'] = "http://localhost/genmon-ch/Data_files/";	// host directory needed by the javaScript script to locate the .json file
/var/www/html/genmon-ch/header.php:		<h1><a href="http://localhost/genmon-ch/index.php"><strong><font color="#FFFFFF">GenMon-CH</font></strong></a></h1>
/var/www/html/genmon-ch/header.php:			$url="http://localhost/genmon-ch";
/var/www/html/genmon-ch/install.html:<span class="co">#change the base url. The code assumes you run on localhost. Might want to change it to your real machine address. If so, run:</span>
/var/www/html/genmon-ch/install.html:<span class="fu">grep</span> -R <span class="st">'localhost'</span> <span class="kw">|</span> <span class="fu">xargs</span> sed -i <span class="st">'s/localhost/yourmachineaddress/g'</span>  
/var/www/html/genmon-ch/map1.js:		"http://localhost:8080/geoserver/cite/wms?",
/var/www/html/genmon-ch/map1.js:			sld: 'http://localhost/genmon-ch/mean_inb_lastgi.xml',
/var/www/html/genmon-ch/map1.js:	var legendbegin="<img src=\"proxy.cgi?url=http://localhost:8080/geoserver/wms%3FREQUEST=GetLegendGraphic%26VERSION=1.0.0%26FORMAT=image/png%26WIDTH=20%26HEIGHT=20%26LAYER=cite:plzo_plz%26sld=http://localhost/genmon-ch/";
/var/www/html/genmon-ch/map1.js:		"http://localhost:8080/geoserver/cite/wms?",
/var/www/html/genmon-ch/map1.js:			sld: 'http://localhost/genmon-ch/max_inb_lastgi.xml',
/var/www/html/genmon-ch/map1.js:		"http://localhost:8080/geoserver/cite/wms?",
/var/www/html/genmon-ch/map1.js:			sld: 'http://localhost/genmon-ch/mean_introgr_lastgi.xml',
/var/www/html/genmon-ch/map1.js:		"http://localhost:8080/geoserver/cite/wms?",
/var/www/html/genmon-ch/map1.js:			sld: 'http://localhost/genmon-ch/num_ind_lastgi.xml',
/var/www/html/genmon-ch/map1.js:		"http://localhost:8080/geoserver/cite/wms?",
/var/www/html/genmon-ch/map1.js:			sld: 'http://localhost/genmon-ch/index_socioec.xml',
grep: /var/www/html/genmon-ch/ol: Is a directory
/var/www/html/genmon-ch/proxy.cgi:                'www.openstreetmap.org', 'sample.avencia.com', 'localhost:8080' ]
```

This result reveals several intersting facts about the internal configuration of the GenMon System. 

1. The main GenMon system consisting of all the php-code is configured such that it listens to port 80 and is served by an apache webserver.
2. The 'geoserver' part listens to port 8080 served by the same apache2 system.
3. For some parts of the 'geoserver', port 8085 is also required.


## Upgrade Singularity
The current version of the singularity container host which comes from the bionic repository is 

```
$ singularity --version
2.6.1-dist
```

We want to upgrade singularity container host to a version past 3.0. The first step to do this is to uninstall the current version. Then we try to install a more current version.

### Step 1: Uninstall
The following command uninstalls the singularity container host

```
sudo apt purge singularity-container
```

We can check with

```
sudo apt list --installed | grep singularity
```

or

```
dpkg --list singularity-container
```

### Step 2: Requirements
In a first step, the requirements listed on '' are checked

```
for p in ca-certificates containernetworking-plugins libc6 libseccomp2 squashfs-tools
do
  echo " * Checking $p ..."
  dpkg --list $p
  sleep 2
done
```

From the output of the above, we have to install 'containernetworking-plugins'. 

```
DWLDIR=/home/quagadmin/source/singularity-container
if [ ! -d "$DWLDIR" ];then mkdir -p $DWLDIR;fi
cd $DWLDIR
wget http://ftp.br.debian.org/debian/pool/main/g/golang-github-containernetworking-plugins/containernetworking-plugins_0.8.6-2_amd64.deb
sudo dpkg --install containernetworking-plugins_0.8.6-2_amd64.deb
#rm -rf containernetworking-plugins_0.8.6-2_amd64.deb
```

### Step 3: Install

The new version is downloaded from 

```
wget http://ftp.br.debian.org/debian/pool/main/s/singularity-container/singularity-container_3.5.2+ds1-1_amd64.deb
sudo dpkg --install singularity-container_3.5.2+ds1-1_amd64.deb
```

After the upgrade the previously created container images could still be run. The only change that had to be done was to add the container directory /var/run/postgresql to the directory binding list.


## Webserver as Service
We use the description under 'https://sylabs.io/guides/3.0/user-guide/running_services.html' for the services and 'https://sylabs.io/guides/3.0/user-guide/networking.html' for the network-arguments as a template. 

The following command pulls and builds the container with an nginx webserver inside.

```
sudo singularity instance start --writable-tmpfs --net --network-args "portmap=8080:80/tcp" docker://nginx web2
```

The container must be started and can then be tested using 'curl'. 

```
sudo singularity exec instance://web2 nginx
$ curl localhost:8080
10.22.0.1 - - [16/Oct/2018:09:34:25 -0400] "GET / HTTP/1.1" 200 612 "-" "curl/7.58.0" "-"
<!DOCTYPE html>
<html>
<head>
<title>Welcome to nginx!</title>
<style>
    body {
        width: 35em;
        margin: 0 auto;
        font-family: Tahoma, Verdana, Arial, sans-serif;
    }
</style>
</head>
<body>
<h1>Welcome to nginx!</h1>
<p>If you see this page, the nginx web server is successfully installed and
working. Further configuration is required.</p>

<p>For online documentation and support please refer to
<a href="http://nginx.org/">nginx.org</a>.<br/>
Commercial support is available at
<a href="http://nginx.com/">nginx.com</a>.</p>

<p><em>Thank you for using nginx.</em></p>
</body>
</html>
```

### Apache 
We use apache2 instead of nginx

```
sudo singularity instance start --writable-tmpfs --net --network-args "portmap=8080:80/tcp" docker://httpd apache2
```

Starting the server

```
sudo singularity exec instance://apache2 httpd
$ curl localhost:8080
<html><body><h1>It works!</h1></body></html>
```


## Changes with Singularity Host Upgrade
The following changes were noticed when going from singularity host version 2.6.2 to 3.5.2. 

### Building Singularity Images
The option '--writable' is no longer available for the 'singularity build' command. The way to create a changeable image is to create a sandbox directory with the '--sandbox' option. An example call is

```
sudo singularity build --sandbox gnm_devel ../../genmon-sidef/def/Singularity_genmon.recipe &> `date +"%Y%m%d%H%M%S"`_gnm.log
```

Once all the changes are done in that directory it can be converted to a '.sif' image file using the following command

```
sudo singularity build `date +"%Y%m%d%H%M%S"`_gnm.sif gnm_devel
```

Then we create the following link

```
SIFIMG=gnm.sif
if [ -e "$SIFIMG" ];then rm $SIFIMG;fi
ln -s `ls -1tr *.sif | tail -1` $SIFIMG
```


## Integration
The findings from above on how to run an apache2 webserver in a singularity instance are now applied to the singularity image that was created before. 

In a first step, the start of a genmon instance is prepared by running 

```
./simg/genmon-sidef/bash/init_gnm_workdir.sh
```

Then the instance with GenMon can be started as

```
sudo singularity instance start  --writable-tmpfs --net --network-args "portmap=8080:8080/tcp" /home/quagadmin/simg/img/genmon/gnm.simg signm
```

The successful start of the instance can be verified using

```
$ sudo singularity instance list
INSTANCE NAME    PID      IP              IMAGE
signm            22091    10.22.0.5       /home/quagadmin/simg/img/genmon/gnm.simg
```

The apache webserver then has to be started with 

```
sudo singularity exec instance://signm httpd
/.singularity.d/actions/exec: 21: exec: httpd: not found
```

In '/.singularity.d/actions/exec' all bash scripts contained in '/.singularity.d/env' are executed. In the 'apache2' instance there is the script '10-docker2singularity.sh' contains the extension of the PATH variable with the path to httpd.

Found the apache2 executable and trying to call it directly

```
sudo singularity exec instance://signm /usr/sbin/apache2
[Thu Aug 20 15:01:13.761430 2020] [core:warn] [pid 16] AH00111: Config variable ${APACHE_RUN_DIR} is not defined
apache2: Syntax error on line 80 of /etc/apache2/apache2.conf: DefaultRuntimeDir must be a valid directory, absolute or relative to ServerRoot
```

Try with based on information in /etc/apache2/apache2.conf

```
sudo singularity exec instance://signm /usr/sbin/apache2ctl start
```

Checking leads to 

```
$ curl localhost:8080
curl: (7) Failed to connect to localhost port 8080: Connection refused
```

For some reasons the wrong port was specified in ports.conf. With 

```
 1838  sudo singularity instance start --writable-tmpfs --net --network-args "portmap=8080:80/tcp" /home/quagadmin/simg/img/genmon/gnm.simg signm
 1839  sudo singularity exec instance://signm /usr/sbin/apache2ctl start
 1840  curl localhost:8080
```

It works.


## Starting New GenMon Image
Using the example from the previous section, we start the instance of the new GenMon Image using

```
sudo singularity instance start  --writable-tmpfs --net --network-args "portmap=90:80/tcp" /home/quagadmin/simg/img/genmon/gnm.sif signm
```

The previous command can be checked with 

```
sudo singularity instance list
```

Starting the apache webserver in the image

```
sudo singularity exec instance://signm /usr/sbin/apache2ctl start
curl localhost
```

The instance is started with multiple portmap arguments

```
sudo singularity instance start  --writable-tmpfs --net --network-args "portmap=90:80/tcp","portmap=8080:8080/tcp" /home/quagadmin/simg/img/genmon/gnm.sif signm
sudo singularity instance list
```

```
sudo singularity exec instance://signm /usr/sbin/apache2ctl start
curl localhost
curl localhost:90
curl localhost:8080
```

Stopping the instance

```
sudo singularity instance stop signm
```

## Restart nginx
The following command can be used to restart nginx

```
sudo systemctl restart nginx
```

The status of the server can be checked with

```
sudo systemctl status nginx
```


## Current Problem

Database log shows

```
Singularity> cat /home/quagadmin/gnm/pglog/20200903171323_postgres.log
2020-09-03 17:13:28.697 CEST [857] LOG:  listening on IPv4 address "127.0.0.1", port 5433
2020-09-03 17:13:28.934 CEST [857] LOG:  listening on Unix socket "/var/run/postgresql/.s.PGSQL.5433"
2020-09-03 17:13:29.049 CEST [858] LOG:  database system was shut down at 2020-09-03 17:13:24 CEST
2020-09-03 17:13:29.101 CEST [857] LOG:  database system is ready to accept connections
2020-09-03 17:13:29.176 CEST [876] FATAL:  role "postgres" does not exist
2020-09-03 17:14:31.404 CEST [881] FATAL:  role "postgres" does not exist
2020-09-03 17:17:46.024 CEST [886] FATAL:  role "root" does not exist
```




## TODOs
This section contains problems and issues which have not been resolved yet.


### Postgresql Port
Check whether the environment variable 'PGPORT' which defines on which port postgresql is listening can be defined in the singularity definition file. According to 'https://sylabs.io/guides/3.0/user-guide/environment_and_metadata.html' this can be done with a special label '%environment' in the singularity definition file. The current running version of 'ssgnm' does not seam to need this. 


### Directory Data_files
Based on several a grep in several php-files, it seams that a subdirectory 'Data_files' is required under '/var/www/html/genmon-ch'. 
